Off the beaten path: \\ ...

### 2018-04-28 paper/gd1.tex
intro refs

% aastex parameters
% \received{not yet; THIS IS A DRAFT}
%\revised{not yet}
%\accepted{not yet}
% % Adds "Submitted to " the arguement.
% \submitjournal{ApJ}
\shorttitle{GD-1 in Gaia DR2}
\shortauthors{price-whelan \& bonaca}

@@

\title{A first look at the GD-1 stellar stream with Gaia DR2}

@@

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold streams
of stars that are extremely valuable tracers of the large- and small-scale
distribution of matter in the Galaxy.
Most of such streams discovered in the Milky Way reside in its halo, and are therefore primarily sensitive to the distribution of dark matter.
We present a sample of highly-probable members of the longest cold stream, GD-1, selected on small parallaxes and retrograde proper motions provided by the \gaia\ mission, and \pans\ photometry consistent with an old and metal-poor population at a distance of $\sim10\,$kpc.
This selection extended the stream by $10^\circ$(?), revealed a possible location of the progenitor and mapped significant density variations along the stream in unprecedented detail.
In addition to several prominent gaps, for the first time we also find evidence of stream members off the main track.
No(?) secular formation mechanisms predict the existence of such spurs, but they are a natural result of a stream interaction with a massive perturber.
Based on the impulse approximation arguments, we estimate that the detected underdensity and the associated spur could have been induced by a perturbation of a $\sim10^7\,\rm M_\odot$(?) object $\sim 0.5\,$Gyr ago.
Our model of GD-1 indicates that the stream crosses the Galactic plane at a large radius, which puts the likely origin of the perturber beyond the stellar disk.
\end{abstract}

@@

Its long length --- and therefore its implied old dynamical age --- and its
particular orbit also make it a prime stream to search for interactions with
dark matter subhalos:
because of its large pericentric distance and retrograde orbit with respect to
the Galactic bar, density variations in the stream are not expected from either
interactions with giant molecular clouds (\citealt{TODO}) or resonant
interactions with the bar (\citealt{Pearson:2017}).
It is particularly exciting, therefore, that deep photometry of the stream has
revealed interesting small-scale ``wiggles'' and density variations along a
$\approx 45^\citc$ span of the stream (\citealt{DeBoer:2017}).
, but color-magnitude only, bin stars]

In this \article, we present a the cleanest view of the GD-1 stream using
astrometric data from the \gaia\ mission data release 2 (\DR).
Combined with photometry from \pans\ ...
We revisit...do not try to model or fit for Galactic potential. Focus on stream properties, data.

>>>>

% aastex parameters
% \received{not yet; THIS IS A DRAFT}
%\revised{not yet}
%\accepted{not yet}
% % Adds "Submitted to " the arguement.
% \submitjournal{ApJ}
\shorttitle{GD-1 in Gaia DR2}
\shortauthors{Price-Whelan \& Bonaca}

@@

\title{Off the beaten path: \\
       ...}
% Gaia DR2 reveals stream stars...
% A first look at the GD-1 stellar stream with Gaia DR2

@@

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold
streams of stars that are extremely valuable tracers of the large- and
small-scale distribution of matter in the Galaxy.
Most stellar streams discovered around the Milky Way reside in the Galactic halo
and are therefore primarily sensitive to the distribution of dark matter.
We present a sample of highly-probable members of the longest cold stream, GD-1,
selected to have small parallaxes and retrograde proper motions using data from
the \gaia\ mission, and \pans\ photometry consistent with an old and metal-poor
population at a distance of $\sim10~\kpc$.
This selection extended the stream by $10^\circ$(?), revealed a possible location of the progenitor and mapped significant density variations along the stream in unprecedented detail.
In addition to several prominent gaps, for the first time we also find evidence of stream members off the main track.
No(?) secular formation mechanisms predict the existence of such spurs, but they are a natural result of a stream interaction with a massive perturber.
Based on the impulse approximation arguments, we estimate that the detected underdensity and the associated spur could have been induced by a perturbation of a $\sim10^7\,\rm M_\odot$(?) object $\sim 0.5\,$Gyr ago.
Our model of GD-1 indicates that the stream crosses the Galactic plane at a large radius, which puts the likely origin of the perturber beyond the stellar disk.
\end{abstract}

@@

Its long length --- and therefore its implied old dynamical age --- combined
with its particular orbit also make it a prime stream to search for interactions
with dark matter subhalos:
because of its large pericentric distance and retrograde orbit with respect to
the Galactic bar, density variations in the stream are not expected from either
interactions with giant molecular clouds (\citealt{Amorisco:2016}) or resonant
encounters with the bar (\citealt{Pearson:2017}).
[Erkal 2016 predict ~1 gap]

From SDSS photometry, hints of gaps and under-densities have been identified in
the GD-1 stream, but with low significance (\citealt{Carlberg:2013}).
More recently, deep photometry of GD-1 from CFHT/Megacam has revealed
interesting small-scale ``wiggles'' and significant density variations along a $\approx 45^\citc$ span of the stream (\citealt{DeBoer:2017}):
[summarize their conclusions].

In this \article, we present clearest view of the GD-1 stream to date using
astrometric data from the \gaia\ mission data release 2 (\DR) combined with
precise photometry from the \pans\ survey.
We find XX under-densities gaps, some consistent with ...
We revisit...do not try to model or fit for Galactic potential.
Focus on stream properties, data.
